Reduced Nucleotide Diversity on a Plant Y Chromosome Following Recent Recombination Suppression

### 2016-08-02 manuscript.tex
Rnotebook, major edits on Ne section

\subsection*{Estimating neutral polymorphism on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2; Table 1). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated such that the true uncertainty in our Y/A diversity estimate obtained by this method is likely underestimated. To address this issue, we used a maximum likelihood approach, implemented in the software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (see Figure S1). We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males. As no significant difference was observed (\X Figure? Table?), we report only results from females.

@@

Such that with no variation in mating success for males or females ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 2 and 3 reduces to the familiar expressions:

@@

respectively. To consider the joint effects of biased sex ratios and variance in male reproductive success (which is more likely in a wind-pollinated annual), our neutrally-expected X/A and Y/A levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{d{f}}=0$) and we considered values for $C^2_{d{m}}$: 0 (no variance), 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3). With the level of neutral polymorphism in a population equalling the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success (assuming equal mutation rates among genes).

>>>>

\subsection*{Estimating polymorphism levels on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2; Table 1). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions of independent loci are likely violated such that the true uncertainty in our Y/A diversity estimate obtained by this method is likely underestimated. To address this issue, we used a maximum likelihood approach, implemented in the software \citep{wright2004hka}, to independently estimate a credibility interval for the Y/A ratio (see Figure S1). We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males. As no significant difference was observed (\X Figure? Table?), we report only results from females.

@@

Such that with no variation in mating success for males or females ($C^2_{d{m}}=C^2_{d{f}}=0$), equations 2 and 3, respectively, reduce to the familiar expressions from \citep{wright1931evolution}:

@@

To consider the joint effects of biased sex ratios and variance in male reproductive success (which is more likely in a wind-pollinated annual), our neutrally-expected X/A and Y/A levels of diversity were calculated assuming no variance in female mating (e.g., $C^2_{d{f}}=0$) and we considered values for $C^2_{d{m}}$: 0 (no variance), 1 (Poisson distribution of offspring), 2 (Poisson*2), and 3 (Poisson*3). With the level of neutral polymorphism in a population equalling the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975,kimura1984}, Figure 1 shows the predicted X/A and Y/A diversity ratios as a function of the sex ratio and variance in male reproductive success (assuming equal mutation rates among genes).
